\section{Info} \label{sec:Info}

This utility prints the system information read from an input structure file,
possibly augmenting it with information from another structure file (\tt{-i[!]
<file>} option) and/or pruning the system to contain only beads from a provided
coordinate file (\tt{-c} option). For the effects of these options, see the
\tt{Examples/Info} folder.

Note that using LAMMPS \data, \ltrj or \xyz input structure file automatically
reads coordinates in those files, but using \vtf does not. This is because a
\vtf timestep does not necessarily contain all beads defined in the \vtf file,
in which case \tt{Info} would automatically prune the system. If coordinates
should be read from the input \vtf structure file (and the system potentially
pruned), just use the \tt{-c} option with that filename.

\tt{Info} can also write this information into a new structure file of the given
format (\tt{-o} option). Allowed output files are \vtsf file, LAMMPS \data file
(either with the extension \tt{.data} or a filename without any recognizable
extension), and DL\_MESO FIELD file (called \field or with the extension
\tt{.FIELD}).

Coordinates are printed to output files that support them, i.e., LAMMPS \data or
\vtf file. These coordinates are 0 if no coordinate file is specified or the
input structure file does not contain them.

There are a few options for the output structure file. For LAMMPS \data file,
\tt{--mass} can be used to specify LAMMPS atom types by their mass, but print
different charges in the \tt{Atoms} section (i.e., \tt{Info}-recognized bead
types which differ only in charge are aggregated into the same LAMMPS atom types
in the \tt{Mass} section). Furthermore, extra atom types (with mass 1) can be
added via the \tt{-ebt} option; these atoms do not appear in the \tt{Atoms}
section of the \data file and can be used as, e.g., extra atom type when srp
potential (i.e., non-crossing bonds created by adding ghost particles between
bonded beads) is used. For \vtsf file, \tt{atom default} bead type can be
specified via the \tt{-def} option; by default, the bead type with the most
unbonded beads is used as the \tt{atom default} type.

\vspace{1em}
\noindent
Usage: \tt{Info <input> <options>}
\noindent
\begin{longtable}{p{0.19\textwidth}p{0.754\textwidth}}
  \toprule
  \multicolumn{2}{l}{Mandatory argument}\\
  \midrule
  \tt{<input>} & input structure file\\
  \midrule
  \multicolumn{2}{l}{Options for input files}\\
  \midrule
  \tt{-i[!] <file>} & extra input structure file; see \tt{Examples/Info}
    directory for the effect of \tt{-i} and \tt{-i!} options (default: None)\\
  \tt{-c <file>}    & input coordinate file (default: None)\\
  \midrule
  \multicolumn{2}{l}{Options for output file}\\
  \midrule
  \tt{-o <file>}    & output coordinate file (default: None)\\
  \tt{-def <bead>}  & only \vtf structure file: use bead type \tt{<bead>} for
    the \enquote{atom default} line (default:bead type with the most unbonded
    beads)\\
  \tt{--mass}       & only \data structure file: define LAMMPS atom types by
    mass but print per-atom charges in the \tt{Atoms} section (default: atom
    types are the same as \tt{Info}-recognized bead types)\\
  \tt{-ebt <int>}   & only \data structure file: number of extra atom types
                      (for, e.g., srp)\\
  \midrule
  \multicolumn{2}{l}{Other options (see the beginning of 
                     Chapter~\ref{chap:Utils})}\\
  \midrule
  \multicolumn{2}{l}{\tt{-st},
                     \tt{-pbc},
                     \tt{--detailed},
                     \tt{--verbose},
                     \tt{--silent},
                     \tt{--help},
                     \tt{--version}}\\
  \bottomrule
\end{longtable}
