\chapter{Format of input/output files}\label{chap:input}

This section describes several file types used by many of the
AnalysisTools utilities. Output files for the utilities themselves are
described in their respective sections in Chapter~\ref{chap:Utils}.

Generally, the utilities work with
\href{https://github.com/olenz/vtfplugin/wiki/VTF-format}{vtf files} which are
convenient text files containing particle coordinates as well as particle
definitions (name, mass, charge, etc.) and molecule connectivity (i.e., bonds
between particles). \TODO Additional information (e.g., bond angles or
dihedrals) can be added from other sources such as input files for
\href{https://www.scd.stfc.ac.uk/Pages/DL_MESO.aspx}{DL\_MESO simulation
package} or \href{https://lammps.sandia.gov/}{LAMMPS molecular dynamics
simulator}.

An \tt{agg} text file describes in an easy-to-read format the compositions of
aggregates. It can be generated via the (\TODO make into refs) \tt{Aggregates}
and \tt{Aggregates-NotSameBeads} utilities.

\section{System description} %{{{
Most of the utilities read system description from the \vtf files
which can be complemented by a DL\_MESO \field or a LAMMPS \data files. Not
all valid \vtf files will work with AnalysisTools; see the following
sections for information on their proper formatting.

The \vtf can either be split into a \vsf structure file
(Section~\ref{ssec:StructureVsf}) and a \vcf coordinate file
(Section~\ref{ssec:CoordinateVcf}), or presented in a single \vtf file
containing both the structure and coordinate sections; note that when individual
\vscf files are mentioned in the manual, they can usually be viewed as \vscf
sections of the combined \vtf file.

In general, the utilities consider only bead types that are present in a \vcf
file, i.e., the beads defined in the \vsf file but missing from the \vcf file
are ignored. The utilities work with orthogonal as well as other paralleloid
boxes (i.e., tri- or monoclinic cells).

\subsection{\vsf structure file} \label{ssec:StructureVsf} %{{{

The structure file contains all information about all beads and bonds (except
for their Cartesian coordinates) in two types of entries. The first type, an
\tt{atom} line, contains a bead definition, while the second one, a \tt{bond}
line, defines which beads are connected.

Each atom line describes a single bead (except for the \tt{atom default}
line, see below) and must follow these rules:

\begin{itemize}
  \item the line starts with \tt{a[tom]} (i.e., \tt{atom} or just \tt{a})
  \item the second string is either a bead index (starting at 0) or the
    \tt{default} keyword; every bead that is not explicitly written in the
    bead definition lines is of the \tt{default} type (in case of multiple
    \tt{default} lines, only the first one is used)
    \begin{itemize}
      \item \TODO allow \tt{atom 0:10 ...} and (possibly) \tt{0:10 ...} lines
    \end{itemize}
  \item the line contains bead name as \tt{n[ame] <char(16)>}
  \item if the bead is in a molecule, the line contains both molecule name
    (\tt{res[name] <char(8)>}) and molecule id (\tt{resid <int>})
    \begin{itemize}
      \item \TODO \tt{res[name]} shouldn't be obligatory; just make up some name
        (e.g., \tt{mol\_1}) if the keyword is missing
    \end{itemize}
  \item \tt{m[ass]}, \tt{charge|q} (i.e., \tt{charge} or \tt{q}), and
    \tt{r[adius]} keywords are read if present
  \item other keywords are ignored
\end{itemize}

Each bond line starts with \tt{b[ond]} followed by two bead indices
(corresponding to their indices from the atom lines) separated by a colon. Note
that the colon must be right after the first index, i.e., no white space between
the number and the colon is allowed.

\tt{AnalysisTools} groups the same beads (and the same molecules) into bead (and
molecule) types. Generally, types of beads are defined solely by their name,
i.e., should two beads share a name but differ in other characteristics, they
will be grouped into the same type with mass, charge, and radius each equal to
that for the bead type's topmost atom line containing the corresponding keyword.
If the keyword is missing from all atom lines, that characteristic is undefined.
The \tt{--detailed} command line option can, however, force some utilities to
distinguish bead types based on all their properties.

Different molecule types, on the other hand, are always distinguished based on
all its characteristics, i.e., name, connectivity, and order of beads (order of
its \vsf indices).

\begin{comment} %{{{
Similarly, different molecule molecules are generally defined based on their
names. If two molecules share a name, but differ in bead order and/or
connectivity, the utility will exit with an error.

However, some utilities can distinguish bead and molecule types based on
other characteristics as well. For example, the \tt{--detailed} option
forces the \tt{Info} utility (\ref{sec:Info}) to consider mass,
charge, and radius along with the name for defining bead types and bead
order and connectivity along with the name for defining molecule types. As
an example, consider the following \vsf file that contains 11 beads and
two molecules:
\end{comment}
 %}}}

As an example, consider the following \vsf file (the extra white space only
aligns bead characteristics into columns) that contains 7 beads and 2
molecules:
\begin{mdframed}
  \begin{tabular}{l@{ }l@{ }l@{ }l@{ }l@{ }l@{ }l}
    atom default & name A & mass 1 & charge 0 & radius 0.5 & &\\
    atom 0       & name B & mass 2 & charge -1 & & resid 1 & resname Mol\\
    atom 2       & name B & mass 1 & charge -1 & radius 1 & resid 2 & resname Mol\\
    atom 4       & name B &        & charge 0.0 & radius 1.5 & &\\
    atom 5       & name A & mass 1 & & & resid 1 & resname Mol\\
    atom 6       & name A & mass 1 & & & resid 2 & resname Mol\\
    bond 0:5&&&&&&\\
    bond 2:6&&&&&&\\
    \end{tabular}
\end{mdframed}
The following table shows how the \tt{--detailed} command line option affects
the identification of bead types:
\begin{center}
\begin{tabular}{lcccccccc}
  \toprule
  \multirow{2}{*}{\parbox{1cm}{\setstretch{0.9}Bead type name}} &
    \multicolumn{4}{c}{\tt{--detailed}} &
    \multicolumn{4}{c}{default behaviour} \\
  \cmidrule(lr){2-5} \cmidrule(lr){6-9}
  & Number & Charge & Mass & Radius & Number & Charge & Mass & Radius\\
  \midrule
  A    & 4      & 0.0    & 1.0  & 0.5    & 4      & 0.0    & 1.0  & 0.5\\
  B    & 1      & -1.0   & 2.0  & N/A    & 3      & -1.0   & 2.0  & 1.0\\
  B\_1 & 1      & -1.0   & 1.0  & 1.0    & ---    & ---    & ---  & ---\\
  B\_2 & 1      & 0.0    & N/A  & 1.5    & ---    & ---    & ---  & ---\\
  \bottomrule
\end{tabular}
\end{center}
Here, the \tt{--detailed} option will also affect identification of molecule
types. Without the option, only one molecule type is detected, but as the
\tt{--detailed} option splits the \tt{B} beads from diferrent molecules into
distinct types, two molecule types are detected (called \tt{Mol} and \tt{Mol_1})

The \vsf file can contain any number of blank lines and comments (lines
beginning with \tt{\#}).

\subsection{\vcf coordinate file} \label{ssec:CoordinateVcf}

The coordinate file contains Cartesian coordinates of the beads and the
size of the paralleloid simulation box. Coordinates are read from a \vcf
file containing either ordered timesteps or indexed timesteps.

An ordered \vcf timestep contains all beads defined in the \vsf file, while an
indexed one can contain only a subset of defined beads. A coordinate line for an
ordered timestep contains only the Cartesian coordinates (where the line order
corresponds to \vsf bead indices), while the one for indexed timestep starts
with a \vsf bead index. Velocities can also be appended to the coordinate lines.
The block of bead coordinates must be continuous without any blank or comment
lines.

The timestep type is defined by a \tt{t[imestep] o[rdered]|i[ndexed]} line
preceding the bead coordinates (i.e., in the timestep preamble). The preamble
also contains simulation box dimensions (\tt{pbc} line); this line is mandatory
for the first timestep in the \vcf file, but optional in the following timesteps
(the box dimensions can be different for each timestep). The \tt{pbc} line can
contain only dimensions (orthogonal box) or dimensions and angles (triclinic
box). The preamble may also contain comment (\tt{\#}-initiated lines) and blank
lines.

The following is a \tt{vcf} file example corresponding to the \vsf example
above:
\begin{mdframed}
  timestep ordered\\
  pbc 1 2 3 \# orthogonal box dimensions\\
  0.0 0.1 0.2 \# ordered coordinate line\\
  0.1 0.2 0.3 0.1 0.0 -0.1 \# coordinate line with velocities\\
  0.2 0.3 0.4\\
  0.3 0.4 0.5\\
  0.4 0.5 0.6\\
  0.5 0.6 0.7\\
  0.6 0.7 0.8\\
  \\
  pbc 1 2 3 100 90 80 \# triclinic box dimensions and angles\\
  timestep indexed\\
  4 0.4 0.5 0.6 \# indexed coordinate line\\
  2 0.2 0.3 0.4
\end{mdframed}
 %}}}

\subsection{DL\_MESO \field\ file} %{{{

\begin{comment}
This file can be used to get mass and/or charge of bead types (only if
missing from the \vsf file) as well as bond parameters and angle and angle
parameters for molecules (these informations are not stored in the \vsf
file).

The format of this file is taken directly from the
\href{https://www.scd.stfc.ac.uk/Pages/DL_MESO.aspx}{DL\_MESO}
software. If the \field file is used only to read mass and/or charge
information about bead types, only the \tt{species} section is required.
This section contains a header line
\begin{verbatim}species <int>\end{verbatim}
where \tt{<int>} is the number of bead types (or species as called by the
DL\_MESO software) in the \field file. Every bead type is then describe by
a single line:
\begin{verbatim}<name> <m> <q> <n>\end{verbatim}
where \tt{<name>} is the bead type name (that must correspond to a bead
name in the \vsf file if its mass/charge is to be read from the \field
file), \tt{m} and \tt{q} are the bead's mass and charge, respectively,
and \tt{<n>} is the number unbonded beads of that type (i.e., beads
not present in a molecule). Not all bead types in the \vsf file must be
present in the \field file. Blank lines are not allowed in the \tt{species}
section.

Should bond and angle information be read as well, a \tt{molecules} section
must follow the \tt{species} section. This section starts with a header line:
\begin{verbatim}molecule <int>\end{verbatim}
where \tt{<int>} is the number of types of molecules. The header is
followed by \tt{<int>} blocks, each describing a single molecule type and
ending with a line:
\begin{verbatim}finish\end{verbatim}
Every molecular block starts with two lines:
\begin{verbatim}<name>
nummols <n>\end{verbatim}
where \tt{<name>} is the molecule's name (that must correspond to a
\tt{resname} in the \vsf file if the bond/angle information is to be read)
and \tt{<n>} is the number of molecules of this type (that does not have to
correspond to the number of \tt{<name>} molecules in the \vsf file).

The rest of the molecule's description is divided into several sub-blocks:
(i) bead order and coordinates, (ii) bond connectivity, and (iii) angles
which is optional. Bead order as well as connectivity must correspond to
that of the \tt{<name>} molecule in the \vsf file.

Block (i) starts with a line:
\begin{verbatim}beads <beads>\end{verbatim}
where \tt{<beads>} is the number of beads in the molecule. Following are
\tt{<beads>} lines for each bead:
\begin{verbatim}<name> <x> <y> <z>\end{verbatim}
where \tt{<name>} is a bead name that must be present in the \tt{<species>}
section and \tt{<x>}, \tt{<y>}, and \tt{<z>} are Cartesian coordinates.

Block (ii) starts with a line:
\begin{verbatim}bonds <n>\end{verbatim}
where \tt{<n>} is the number of bonds in the molecule. Following are
\tt{<n>} lines for each bond:
\begin{verbatim}harm <l> <m> <k> <r>\end{verbatim}
where \tt{<l>} and \tt{<m>} are bead indices of connected beads that run
from 1 to \tt{<beads>} and \tt{<k>} and \tt{<r>} are the strength and
equilibrium distance, respectively, of a harmonic oscillator (AnalysisTools
assumes bonds use harmonic potential).

Block (iii) starts with a line:
\begin{verbatim}angles <n>\end{verbatim}
where \tt{<n>} is the number of angles in the molecule. Following are
\tt{<n>} lines for each angle:
\begin{verbatim}harm <l> <m> <n> <k> <alpha>\end{verbatim}
where \tt{<l>}, \tt{<m>}, and \tt{<n>} are bead indices defining the angle
that run from 1 to \tt{<beads>} and \tt{<k>} and \tt{<alpha>} are the
strength and equilibrium angle, respectively, of a harmonic oscillator
(AnalysisTools assumes angles use harmonic potential). This block is optional.

Anything between the last angle line and the \tt{finish} keyword is ignored.

Several examples of a \field file can be found in the \tt{Examples}
directory, e.g., in \tt{Examples/AddToSystem}.
\end{comment}
 %}}}

\subsection{LAMMPS \data\ file} %{{{
 %}}}
  %}}}

\section{Aggregate file (\tt{agg})} \label{ssec:AggFile} %{{{

A \tt{file.agg} is generated using
\tt{Aggregates} (or \tt{Aggregates-NotSameBeads}) utility. The
file contains information about the number of aggregates in each timestep
and which molecules and monomeric (i.e., unbonded) beads belong to which
aggregate. It serves as an additional input file for utilities that
calculate aggregate properties; \tt{agg} file is, therefore, linked to
the \vcf file that was used to generate it.

The \tt{agg} file is a simple text file. The first line contains the
command used to generate it -- parts of this command may be necessary for
subsequent analysis of aggregates. The second line is blank, and from the
third line, the data for individual timesteps are shown. It follows these
rules:

\begin{itemize}[topsep=0pt,itemsep=0pt]
  \item each timestep starts with \tt{Step: <int>} (only \tt{Step}
    keyword is read by the utilities)
  \item the second line is the number of aggregates in the given
    timestep and is followed by a blank line
  \item there are two lines for each aggregate:
  \begin{enumerate}[topsep=0pt,itemsep=0pt]
    \item number of molecules in the aggregate followed by their indices
      taken from the \vsf file (\tt{resid} indices)
    \item number of monomeric beads in the aggregate followed by their
      indices taken from the \vsf file (\tt{atom} indices)
  \end{enumerate}
  \item no blank or comment lines are allowed inside the aggregate block
  \item not all molecules present in the \vcf file used to generate
    this file must be present in every timestep
\end{itemize}

Note that the term aggregate also refers to free chains (i.e., fully
dissolved chains).

When the keyword \tt{Last} is present instead of \tt{Step}, it
signalises the end of the \tt{agg} file; no utility will read anything
beyond this keyword.

Besides using this file for further analysis by other utilities, the
indices can be used in \href{http://www.ks.uiuc.edu/Research/vmd/}{vmd} to
visualize, e.g., only a specific aggregate by using \tt{resid 1 2 3} in the
\tt{Selected Atoms} box in vmd.

An example of an \tt{agg} file can be found in the \tt{Examples/DistrAgg}
directory. %}}}
