\section{JoinSystems} \label{sec:JoinSystems}

This utility takes an existing system and adds new beads read from a \field into
it, placing them either completely randomly or according to supplied
constraints.

There are two types of constraints which can be combined: place new beads (i)
specified distance from other beads or (ii) in a specified interval in x-, y-,
and/or z-axi directions. In (i), options \tt{-ld} and/or \tt{-hd} specify the
distance; if present, these must be accompanied by \tt{-bt} option.  Then, the
new beads are placed at least \tt{-ld <float>} and at most \tt{-hd <float>}
distance from beads specified by the \tt{-bt} option. In (ii), options \tt{-cx},
\tt{-cy}, and \tt{-cz} basically change the box size for the added beads; for
example, (\TODO real vs. relative units) \tt{-cx 5 9} would generate x
coordinates only in the interval $\langle5,9)$. All these options can be
combined, but note that \tt{AddRandomToSystem} does not perform any sanity
checks, that is, if any combination of the provided options is impossible to
achieve, the utility will run forever.

\input{JoinSystems-tikz}

% \input{JoinSystems-tikz-a}
% \input{JoinSystems-tikz-b}
% \input{JoinSystems-tikz-c}

For added molecules, either the molecule's first bead (\TODO whichs default?
default behaviour) or its geometric centre (\tt{--gc} option\TODO{implement})
obey the constraints. The coordinates of the remaining beads in the molecule are
governed by the coordinates in the \tt{FIELD} file. Therefore, not all the
molecular beads necessarily obey the constraining rules. Molecules are added
with a random orientation unless the \tt{--no-rotate} switch is used.

\TODO I guess that's not true? All added beads are at the end, no?
The new unbonded beads appear in the resulting files after
the unbonded beads from the original system (but before any bonded beads).
If molecules are added, \tt{AddRandomToSystem} places them at the end.

\TODO Not true anymore---the opposite is true, in fact (\tt{--add} switch)!
By default, the new beads are added to the system, increasing the total
number of beads. However, if \tt{--switch} is used, the new beads
replace beads in the original system. The replaced beads are either neutral
beads with the lowest indices (as per ordering in the provided structure
file) or beads of the type specified using \tt{-xb} option. In both
cases, the beads to be replaced must be unbonded.

The size of the new simulation box can be changed using the \tt{-b} option.
There are no constraints on the size of the cuboid box; if the new box is small,
not all beads from the original system are necesserily inside the box (although
the added once always are). (\TODO does \tt{--centre} makes sense? I don't think
so. Or do I? What about using a pure bilayer box, ensuring the bilayer is in the
box centre? The origin of the new box coincides with the origin of the original
box unless \tt{--centre} option is used; in that case, the centres of the two
boxes coincide.)

Several examples of using the utility are provided in the
\tt{Examples/AddRandomToSystem} directory.

\vspace{1em}
\noindent
Usage: \tt{AddToSystem <input.vcf> <out.vsf> <out.vcf> <options>}
\noindent
\begin{longtable}{p{0.235\textwidth}p{0.709\textwidth}}
  \toprule
  \multicolumn{2}{l}{Mandatory arguments} \\
  \midrule
  \tt{<input.vcf>} & input coordinate file (either \tt{vcf} or
    \tt{vtf} format) \\
  \tt{<out.vsf>} & output \tt{vsf} structure file for the new
    system \\
  \tt{<out.vcf>} & output \tt{vcf} coordinate file for the new
    system \\
  \toprule
  \multicolumn{2}{l}{Non-standard options} \\
  \midrule
  \tt{-st <int>} & \tt{<input.vcf>} timestep to add new beads
    to (default: 1) \\
  \tt{-xyz <name>} & also save coordinates to \tt{xyz} file \\
  \tt{--switch} & replace original beads instead of increasing the
    total number of beads \\
  \tt{-b <x> <y> <z>} & side lengths of the new simulation box \\
  \tt{--centre} & place the original simulation box in the middle of
    the new one \\
  \midrule
  \multicolumn{2}{c}{Random placement} \\
  \midrule
  \tt{-f <name>} & \tt{FIELD} file specifying additions to the
    system (default: \tt{FIELD}) \\
  \tt{-ld <float>} & lowest distance from beads specified by
    \tt{-bt} option \\
  \tt{-hd <float>} & highest distance from beads specified by
    \tt{-bt} option \\
  \tt{-bt <bead names>} & bead types to use in conjunction with
    \tt{-ld} and/or \tt{-hd} options \\
  \tt{-cx <num> <num2>} & constrain x coordinate to interval
    $\langle num, num2)$ \\
  \tt{-cy <num> <num2>} & constrain y coordinate to interval
    $\langle num, num2)$ \\
  \tt{-cz <num> <num2>} & constrain z coordinate to interval
    $\langle num, num2)$ \\
  \tt{--gc} & use molecule's geometric centre for distance check \\
  \midrule
  \multicolumn{2}{c}{Provided coordinates} \\
  \midrule
  \tt{-vtf <vsf> <vcf>} & add ready-made system from \tt{vtf} files
    instead of randomly placing beads read from the \tt{FIELD} file \\
  \bottomrule
\end{longtable}
