\section{SelectedVcf} \label{sec:SelectedVcf}

This utility creates a new coordinate file containing only beads of specified
types; any of the \xyz, \vcf, or \tt{lammpstrj} formats can be used as output.

Using \tt{--reverse} switch, the bead types can also be specified in reverse,
i.e., bead types to be excluded from the output file(s).

Which timesteps are saved can be controlled using several options: \tt{-st}
option for the first timestep to save and \tt{-e} option for the last timestep
to save; \tt{-sk} for the number of ignored steps per one saved (i.e., \tt{-sk
9} means every 10th timestep is saved). On the other hand, to explicitly specify
which steps to save, the \tt{-n} option is used; SelectedVcf then ignores the
\tt{-st}, \tt{-e}, and \tt{-sk} options if present. A maximum of 100 steps can
be specified for the \tt{-n} option. Finally, using \tt{--last} option saves
only the last step from the input coordinate file (all the previous options are
ignored if \tt{--last} is used).

There is also an option to remove periodic boundary conditions for molecules
(i.e., to join them) via \tt{--join} switch. Conversely, the simulation box can
be wrapped (i.e., the periodic boundary conditions applied, putting all beads
inside the box) via \tt{--wrap} switch. If both and \tt{--wrap} and \tt{--join}
options are used, the simulation box is first wrapped and then the molecules are
joined.

Usage:

\vspace{1em}
\noindent
\tt{SelectedVcf <input> <output> <bead type(s)> <options>}

\vspace{1em}
\noindent
\begin{longtable}{p{0.22\textwidth}p{0.724\textwidth}}
  \toprule
  \multicolumn{2}{l}{Mandatory arguments}\\
  \midrule
  \tt{<input>}        & input coordinate file\\
  \tt{<output>}       & output coordinate file\\
  \tt{<bead type(s)>} & bead type names to save (can be omitted if
    \tt{--reverse} is used)\\
  \midrule
  \multicolumn{2}{l}{Options}\\
  \midrule
  \tt{--reverse} & reverse bead types to save, i.e., exclude \tt{<bead type(s)>}
    instead of including them; if no \tt{<bead type(s)>} are specified, all bead
    types are saved\\
  \tt{--join}    & join molecules by removing periodic boundary conditions\\
  \tt{--wrap}    & wrap simulation box (i.e., apply periodic boundary
    conditions)\\
  % \tt{-st <n>}   & starting timestep for calculation (default: 1)\\
  % \tt{-e <n>}    & ending timestep for calculation (default: none)\\
  % \tt{-sk <n>}   & number of steps skip per one used (default: 0)\\
  \tt{-n <int(s)>} & save only specified timesteps\\
  \tt{--last} & save only the last step\\
  \midrule
  \multicolumn{2}{l}{Other options (see the beginning of 
                     Chapter~\ref{chap:Utils})}\\
  \midrule
  \multicolumn{2}{l}{\tt{-st},
                     \tt{-e},
                     \tt{-sk},
                     \tt{-i},
                     \tt{--variable},
                     \tt{-pbc},
                     \tt{--detailed},
                     \tt{--verbose},
                     \tt{--silent},}\\
  \multicolumn{2}{l}{\tt{--help},
  \tt{--version}}\\
  \bottomrule
\end{longtable}
