\chapter{Utilities} \label{chap:Utils}

All utilities have command line options affecting their behaviour. Most of the
utilities share several standard command line options.

The following options are for all utilities and are not described with the
utilities themselves.\\[1em]
\noindent
\begin{tabular}{p{0.14\textwidth}p{0.804\textwidth}}
  \toprule
  \multicolumn{2}{l}{General options common to all utilities}\\
  \midrule
  \tt{--verbose} & print information about all bead and molecule types to the
    screen\\
  \tt{--silent}  & run silently, i.e., without any output at all (overrides
    \tt{--verbose} option)\\
  \tt{--help}    & print short description of the utility and usage\\
  \tt{--version} & print version of the utilities and exit\\
  \bottomrule
\end{tabular}\\[1em]

The following options are used by most utilities, and because they affect their
behaviour, they are also listed in the utilities' descriptions.\\[1em]
\noindent
\begin{tabular}{p{0.14\textwidth}p{0.804\textwidth}}
  \toprule
  \multicolumn{2}{l}{Options for structure and coordinate input files}\\
  \midrule
  \tt{-i <name>}  & use specified structure file, see
    Section~\ref{sec:StructFile} for possible formats\\
  \tt{--detailed} & only \vtf structure file: detailed recognition for bead
    (and, consequently, molecule) types based on the bead names as well as
    charges, masses, radii \\
  \tt{--variable} & only \vtf coordinate file: assume
    indexed timesteps with variable numbers of beads (otherwise, the same number
    of beads in each timestep is required)\\
  \tt{-pbc <n>}   & only \xyz file: position of the first box dimension in
    the file's comment line at the beginning of each timestep\\
  \tt{-st <n>}    & any coordinate file: start calculations at \tt{n}-th
    timestep\\
  \tt{-e <n>}     & any coordinate file: end calculations at \tt{n}-th
    timestep\\
  \tt{-sk <n>}    & any coordinate file: skip every \tt{n} timesteps during
    calculations\\
  \bottomrule
\end{tabular}\\[1em]

For \vtf structure files, the bead types are defines solely based on their name;
the other properties (mass, charge, and radius) are taken from the first
\tt{a[tom]} line with the bead of that name that contains the appropriate
keyword. For example, lines

\noindent
\tt{atom default name A m 1}\\
\tt{atom 1 name B}\\
\tt{atom 2 name B m 2}\\
\tt{atom 3 name A m 2}\\
\tt{atom 4 name A}\\[-1em]

\noindent
would define two bead type called \tt{A} and \tt{B} with masses equal to 1 and
2, respectively. Using the \tt{--detailed} option would split the \tt{A} beads
into three types: \tt{A} and \tt{A_1} with masses equal to 1 and 2,
respectively, and \tt{A_2} with undefined mass (it remains undefined because
there is an ambiguity---should the last bead have mass 1 or 2?). On the other
hand, the \tt{B} beads would be assigned the same bead type with mass equal to 2
(there is no ambiguity because only one mass is specified for the \tt{B} beads).
See \tt{Examples/Info} directory for a more complete example.

For \vtf coordinate file with indexed timesteps, all timesteps are expected to
have the same number of beads (though not necessarily the same beads). If some
timesteps contain different number of beads, \tt{--variable} option must be
used. The option is implemented because without knowing in advance the number of
beads in a timestep, reading that timestep is slower (significantly so for large
systems).

Plain \xyz format provides only bead names and coordinates, but the comment line
at the beginning of every timestep can store extra information.
Here, only box dimensions are read, but because there are extended \xyz formats
containing more information, the \tt{-pbc} option is used to specify where the
box dimensions are (as well as to tell the utility there are box dimensions at
all). For example, \tt{-pbc 3} must be specified for the following comment line

\tt{something more 10 10 10 90 60 30}.

\noindent
This would define a paralleloid box with $a=b=c=10$ and angles
$\alpha=90^\circ$, $\beta=60^\circ$, and $\gamma=30^\circ$. If less than 6
numbers were provided, only the first three would be used to define a cuboid
box.

\input{AddRandomToSystem}
% \input{Aggregates}
% \input{AngleMolecules}
\input{Average}
% \input{BondLength}
% \input{Config}
% \input{DensityAggregates}
\input{DensityBox}
% \input{DensityMolecules}
% \input{DihedralMolecules}
% \input{DistrAgg}
% \input{GenLayers}
% \input{GenSystem}
% \input{GyrationAggregates}
% \input{GyrationMolecules}
\input{Info}
% \input{JoinAggregates}
% \input{JoinRuns}
% \input{lmp_data}
% \input{lmp_vtf}
\input{Local}
% \input{Orientation}
% \input{PairCorrel}
% \input{PotentialAggregates}
\input{SelectedVcf}
% \input{Surface}
% \input{traject}
% \input{TransformVsf}
% \input{VisualizeAgg}
